For submission to the Journal of Integrated Pest Management

### 2021-05-21 ofv_fife_2021.Rmd
Added information on OFV strain

We describe the first outbreaks of *Orchid fleck dichorhavirus*, belonging to the orchid-infecting subgroup (OFV-Orc), from three unreported hosts: *Liriope muscari* (Decaisne) Bailey, *Ophiopogon intermedius* Don and *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea) in Leon and Alachua Counties, FL. Strains of OFV-Orc infect over 50 plant species belonging to the plant families Orchidaceae, Asparagaceae (Nolinoidaea), and infects *Citrus* (Rutaceae), as citrus leprosis disease. The only known vectors of OFV-Orc are flat mites from the genus *Brevipalpus* (Trombidiformes: Tenuipalpidae). Florida has various native and introduced plants in the landscape which *Brevipalpus* spp. feed on, which are potentially susceptible to OFV-Orc.

Chlorotic ringspots and flecking were seen affecting Liriopogons (*Liriope* spp. and *Ophiopogon* spp.) in Leon County, FL. Adjacent *A. elatior* also appeared chlorotic. Local diagnostics returned negative for common plant pathogens, therefore samples were sent to the Florida Department of Agriculture and Consumer Services for identification. OFV-Orc was confirmed in *L. muscari* and *A. elatior* via conventional reverse transcription polymerase chain reaction and Sanger sequencing of amplicons. Amplicons had a 98% identity with known sequences of Orchid fleck virus available in NCBI GenBank.

New samples from symptomatic plants were collected in both Leon and Alachua counties and sent to the USDA-ARS. Conventional RT-PCR, RT-qPCR and Sanger sequencing confirmed the presence of Orchid fleck virus, OFV-Orc subgroup. High-throughput sequencing detected OFV-Orc and OFV-Orc2 in *A. elatior* and *L. muscari*, respectively. Three potential mite vectors were identified via cryo-scanning electron microscopy (Cryo-SEM): *Brevipalpus californicus* (Banks) sensu lato, *B. obovatus* Donnadieu, and *B. confusus* Baker.

@@

The initial site of collection was visited two more times during 2020 to gather plants for identification of the unidentified plant pathogen. These surveys were conducted during July and August to collect more putatively infected plants, including more *Liriope* spp. as well as a new species which also belongs to the family Asparagaceae; *Aspidistra elatior* Blume, which was suspected to be infected due to unusually chlorotic leaves (`r citef("fig_2")`). Upon collection, these new samples were sent to the Florida Department of Agriculture and Consumer Services (FDACS) for identification. 
The FDACS determined that the virus was *Orchid fleck dichorhavirus* using previously published primers and methods [@Kubo2006a; @Kubo2006b; @Kubo2009a; @Kubo2009b; @RamosGonzalez2015]. Orchid subgroup OFV-Orc was identified following the methods described in @Kondo2017.

Samples sent to the USDA confirmed the presence of OFV via one step conventional RT-PCR and RT-qPCR, using generic R2-Dicho-GF and R2-Dicho-GR primers, amplifying ~800 nt of L-gene (RNA2) amplicon [@Roy2020]. Follow-up Sanger sequencing demonstrated a shared 98\% nucleotide identity with the orchid strain subgroup, OFV-Orc (isolates So and Br with GenBank Accession numbers: AB244418 and MK522807, respectively) [@Kondo2006; @Kondo2017]. High Throughput sequencing results confirmed the presence of OFV-Orc2 strain in the *Liriope* sample.

>>>>

We describe the first outbreaks of *Orchid fleck dichorhavirus*, belonging to the orchid-infecting subgroup (OFV-Orc), from three unreported hosts: *Liriope muscari*, cv. 'Gigantea (Decaisne) Bailey, *Ophiopogon intermedius* Don and *Aspidistra elatior* Blume (Asparagaceae: Nolinoidaea) in Leon and Alachua Counties, FL. Strains of OFV-Orc infect over 50 plant species belonging to the plant families Orchidaceae, Asparagaceae (Nolinoidaea), and infects *Citrus* (Rutaceae) as citrus leprosis disease. The only known vectors of OFV-Orc are flat mites, genus *Brevipalpus* (Trombidiformes: Tenuipalpidae). Florida has various plants in the landscape which *Brevipalpus* spp. feed on, which are potentially susceptible to OFV-Orc.

Chlorotic ringspots and flecking were seen affecting Liriopogons (*Liriope* spp. and *Ophiopogon* spp.) in Leon County, FL. Adjacent *A. elatior* also appeared chlorotic. Local diagnostics returned negative for common plant pathogens, therefore samples were sent to the Florida Department of Agriculture and Consumer Services and USDA-ARS for identification.

OFV-Orc was confirmed in *L. muscari* and *A. elatior* via conventional and quantitative reverse transcription polymerase chain reaction. Sanger sequencing of amplicons found 98% nucleotide identity with orchid-infecting strains of Orchid fleck virus available in NCBI GenBank. High-throughput sequencing detected OFV-Orc and OFV-Orc2 in *A. elatior* and *L. muscari*, respectively.

Three potential mite vectors were identified via cryo-scanning electron microscopy (Cryo-SEM): *Brevipalpus californicus* (Banks) sensu lato, *B. obovatus* Donnadieu, and *B. confusus* Baker.

@@

The initial site of collection was visited two more times during 2020 to gather plants for identification of the unidentified plant pathogen. These surveys were conducted during July and August to collect more putatively infected plants, including more *Liriope* spp. as well as a new species which also belongs to the family Asparagaceae; *Aspidistra elatior* Blume, which was suspected to be infected due to unusually chlorotic leaves (`r citef("fig_2")`). Upon collection, these new samples were sent to the Florida Department of Agriculture and Consumer Services (FDACS) for identification. 
The FDACS determined that the virus was *Orchid fleck dichorhavirus* using previously published primers and methods [@Kubo2006a; @Kubo2006b; @Kubo2009a; @Kubo2009b; @RamosGonzalez2015]. Orchid subgroup 1, OFV-Orc was identified following the methods described in @Kondo2017.

Samples sent to the USDA confirmed the presence of OFV via one step conventional RT-PCR and RT-qPCR, using generic R2-Dicho-GF and R2-Dicho-GR primers, amplifying ~800 nt of L-gene (RNA2) amplicon [@Roy2020]. Sanger sequencing demonstrated a shared 98\% nucleotide identity with the orchid strain subgroup, OFV-Orc (isolates So and Br with GenBank Accession numbers: AB244418 and MK522807, respectively) [@Kondo2006; @Kondo2017]. High Throughput sequencing results confirmed the presence of OFV-Orc2 strain in the *Liriope* sample.
